---
title: "hw6"
author: "Mike Lehman"
date: "October 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## #1

```{r}
library(ggplot2)
data(diamonds)

price_df <- as.data.frame(diamonds[,"price"])
colnames(price_df) <- c("price")
summary(price_df$price)
d <- ggplot(price_df["price"], aes(x = price))
d + geom_density()
```

## #2

```{r}
d_log10 <- ggplot(price_df["price"], aes(x = log(price, base = 10)))
summary(log(price_df$price, base = 10))
d_log10 + geom_density()
```

Performing a log (base 10) transformation on the price variable was worthwhile in this case because it reduced the scope of the varialbe's distribution and brought the mean and median much closer than they were in the raw data.

Since the range of the raw data for the price variable was so wide and all values were positive, a log transformation was worthwhile.


## #3

```{r}
set.seed(123)
train_sample <- sample(53940, 26970)

diamonds_train <- diamonds[train_sample, ]
diamonds_test  <- diamonds[-train_sample, ]

str(diamonds_train)
str(diamonds_test)
```

## #4

```{r}
price_reg <- lm(price ~ carat + cut + color + clarity, data = diamonds_train)
price_reg

price_reg_log10 <- lm(log(price, base = 10) ~ carat + cut + color + clarity, data = diamonds_train)
price_reg_log10
```

## #5

```{r}
# create log base(10) transformed version of price feature for reference
diamonds_test$price_log10 <- log(diamonds_test$price, base = 10)
diamonds_test$PricePrediction <- predict(price_reg, newdata = diamonds_test)
diamonds_test$PricePrediction_log10 <- predict(price_reg_log10, newdata = diamonds_test)
head(diamonds_test)
```

## #6

```{r}
a <- ggplot(data = diamonds_test, aes(x = PricePrediction, y = PricePrediction - price))
a + geom_point(alpha = 0.2, color = "black") + geom_smooth(aes(x = PricePrediction, y = PricePrediction - price), color = "black")

b <- ggplot(data = diamonds_test, aes(x = PricePrediction_log10, y = PricePrediction_log10 - price_log10))
b + geom_point(alpha = 0.2, color = "black") + geom_smooth(aes(x = PricePrediction_log10, y = PricePrediction_log10 - price_log10), color = "black")
```

## #7

```{r}
confint(price_reg)
par(mfrow=c(2,2))
plot(price_reg)

c <- ggplot(data = diamonds_test, aes(x = PricePrediction, y = price))
c + geom_point(alpha = 0.2, color = "black") + geom_smooth(aes(x = PricePrediction, y = price), color = "black") + geom_line(aes(x = log(PricePrediction, base = 10), y = log(price, base = 10)), color = "blue", linetype = 2)

confint(price_reg_log10)
par(mfrow=c(2,2))
plot(price_reg_log10)

d <- ggplot(data = diamonds_test, aes(x = PricePrediction_log10, y = price_log10))
d + geom_point(alpha = 0.2, color = "black") + geom_smooth(aes(x = PricePrediction_log10, y = price_log10), color = "black")
```

From the above graphs and plots, we can see that both models do an acceptable job of predicting the price. 

Looking at the confidence intervals for each model, it appears that carat, cut, and clarity are the most strongly correlated with the dependent variable, price. Both models also seem to have a somewhat linear pattern. There is not perfect linearity here, but the lines in the above graphs show general linearity and the values are largely consistent among the scatter.

If choosing which model is more accurate, the log(base 10) transformed model yielded better results between the two. This is likely because transforming the price variable brought the range of values closer together, and the median a better resprentative of a specific candidate.