---
title: "IS677 HW13"
author: "Mike Lehman"
date: "November 28, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(foreach)
library(bigmemory)
library(biganalytics)
```

```{r}
#1
airline_08 <- read.big.matrix("2008.csv", type = "integer", header = TRUE, backingfile = "2008.bin", descriptorfile = "2008.desc")
```

```{r}
#2
options(scipen = 999, digits = 0)

start_time <- Sys.time()

summary(airline_08)

stop_time <- Sys.time() - start_time
stop_time
```

Machine Details:
Surface Pro 4
Windows 10
16.0 GB RAM
2 physical cores, 4 threads

```{r}
#3
late_flights_ind <- mwhich(airline_08, "ArrDelay", 0, "gt")
nrow(airline_08[late_flights_ind,])
```

```{r}
#4
early_flights_ind <- mwhich(airline_08, "ArrDelay", 0, "lt")
nrow(airline_08[early_flights_ind,])
```

```{r}
#5
ontime_flights_ind <- mwhich(airline_08, "ArrDelay", 0, "eq")
nrow(airline_08[ontime_flights_ind,])
```

```{r}
#6
distances <- airline_08[,"Distance"]
mean(distances)
```

```{r}
#7
on_time_distances <- airline_08[ontime_flights_ind,"Distance"]
mean(on_time_distances)
```

```{r}
#8
early_distances <- airline_08[early_flights_ind,"Distance"]
mean(early_distances)
```

```{r}
#9
late_distances <- airline_08[late_flights_ind,"Distance"]
mean(late_distances)
```

#10
This question cannot be answered because the read.big.matrix command can only accept one data type for all the values it imports from an external file. We used the integer data type as most of the data points were integers/numeric in nature. Subsequently, the bigmemory package imported the Dest feature with all NAs as the original value was of data type string since they are the corresponding IATA airport code. 